\section{Battle}
\par
Battles are fought as a result of Event, Political Action and Movement card play. Each battle must be completed before fighting the next battle. Battles occurring as the result of Event and Political Action card play are resolved immediately with the order determined by the active player. The player who selected a movement action and moved first this round determines the order in which battles are fought as the result of a movement card play.

In any area in which there is a battle both players reveal their units by tipping their blocks forward so that they show their current strength. After each battle is completed, return any surviving units upright before proceeding to the next battle. After all units have taken one battle action, one battle round is considered to have passed. Repeat the sequence for a 2nd and 3rd round if necessary.

Most battles are fought for a maximum of 3 battle rounds, though when crossing the Rhine into Germania or amphibiously invading an area there are only 2 battle rounds. The attacker must retreat if the defender is not eliminated and does not retreat at the end of the maximum number of battle rounds. This procedure is repeated until all battles are complete.

Each unit may fire, retreat or pass during each battle round. This is called a battle action. The order in which a unit resolves its battle action depends on its initiative. Any defending units with an "B" initiative take their battle action first, followed by all attacking "B" units. Then all defending "C" units take their action, followed by all attacking "C" units. Finally, all defending "D" units take their action, followed by all attacking "D" units.

\textbf{Note:} If the Caesar unit (Legion X) is involved in a battle, it always takes its action first, regardless of enemy initiative or circumstances, since it is the only "A" rated block.

Each unit may “fire” by rolling as many dice as its strength. A hit is scored for every roll less than or equal to its battle rating. Against most units each hit reduces an enemy unit by one strength point. This is indicated by rotating the unit counter-clockwise for each hit. Enemy units may not be targeted individually. Each hit is applied to the strongest enemy unit. If two or more enemy units are the same strength, the owner decides on which unit to apply the hit. Units eliminated in battle are set aside until the End of Turn Phase.

\textbf{Exceptions:} In mixed forces, Roman or German units must take a hit before allied Barbarian units if they have the same strength.

Some units require two hits to be reduced by one strength. If possible, all hits must be applied. If an odd number of hits remain at the end of a battle round after applying all hits, the excess are ignored, except in sieges. See below.

\textbf{A. Reserves}
\par
Multiple groups may attack or defend the same area, moving across the same or different borders. However, only one group is considered the main group. All other groups are considered reserves. In the case of defending units, the units that began the action in an area are considered the main group. Any units that moved to help defend the area are considered reserves. In the case of attacking units, the players attacking an area decides which group is the main group and which groups are reserves.

Reserve units may not perform a battle action or suffer hits, nor are they revealed, in the first round of battle. They are revealed and participate in the 2nd round of battle, even if all other friendly units have been eliminated. If the attacking player eliminates all defending units before defending reserve units arrive, then the original attacker is now considered the defender, and the original defender is now considered the attacker. Battle is then resolved normally.

\textbf{B. Sieges}
\par
Defending units that start a battle in an area with a fortress may declare that they are inside the fortress, up to the limit of the supply value of the fortress. Thus, most fortresses may only hold one unit, while Transalpine Gaul and Avaricum can hold up to two units. Defending reserve units do not have this option.

If a Gallic unit is defending in its home area with a fortress, then that unit must be chosen first over other potential units to defend inside a fortress, if any are chosen at all. If there is more than one Gallic unit defending its home area, then the owning player may choose which one goes inside the fortress, and the other unit must fight outside the fortress.

In the circumstance where there are defending units both outside and inside a fortress, a battle is fought normally against the outside unit(s) first. If the attackers successfully defeat the outside units within 3 battle rounds, either by eliminating them or forcing them to retreat, then the attacker has the option of either assaulting the fortress or retreating. If retreat is chosen at this point, it happens prior to the defender rolling any dice.

If the attacker chooses to assault, then the attacker gets a fresh set of 3 battle rounds to prosecute the assault. If the assault is not successfully completed within the allotted number of rounds, then the attacker must retreat.

Units defending in a fortress against an assault are treated as having one better initiative than their current rating, i.e. a unit that has a "C" initiative is treated as a "B", a unit that has a "B" initiative is treated as an "A". Any units already rated an "A" are unaffected.

In addition, the defending unit has double defense, i.e. it requires two hits to score one point of damage. You can mark any "half hits" by turning the block 45 degrees. These half hits are retained between battle rounds, but are discarded once the assault is over. For forts that can hold more than one unit, any half hit must be applied to a unit that has already suffered a half hit, if any.

Defending units inside a fortress may not retreat or reorganize after battle unless the attacking units retreated or were eliminated first.

\textit{Designer's Note: Note that I deliberately do not use a "siege point" system that is popular with other ancients games. Historically the Roman sieges of the various Gallic fortified towns were very brief, and were usually resolved by the Romans cutting off their enemy's food and water supply. Once that was accomplished, the barbarians would soon surrender.}

\textit{The largest and longest sieges of the campaign at Avaricum and Alesia lasted just a few weeks, but most seem to have been successfully prosecuted within a few days. For this reason I do not provide siege rules that are more detailed because they simply wouldn't be appropriate at this scale.}

\textbf{B2. Reserve Units Entering a Besieged Area}

If reserve units are entering an area where the defenders have retreated into a fort, the siege is temporarily interrupted and a field battle occurs, with the besieging units considered the defender. This battle may last up to two rounds, maximum.

Any besieged units may optionally sortie out to join their Reserve forces in this situation, but if they elect to retreat then they may only retreat back into the fortress. Any half hits that were scored before the reserve forces arrived are retained.

Once resolved, the sieging units may proceed with their remaining two siege rounds, assuming they win the field battle.

\textit{Example: On the Roman player's turn, four Roman legions have entered the Mandubii home area. Vercingetorix is the only unit there, with the Mandubii and Senones units having retreated to an adjacent area earlier in the turn. Vercingetorix retreats into the fortress (Alesia).}

\textit{The Romans conduct a siege on the first round, scoring 2 hits for one point of damage. However, the Barbarian player has moved four Gallic units into the area from adjacent areas to try and save Vercingetorix, making them Reserve units entering a besieged area. Consequently, the Roman legions are now defending units against attacking Barbarian units in a field battle. The Barbarian player elects to sortie out Vercingetorix to join the fight.}

\textit{Unfortunately for the Barbarian player, the attack does not go well, and after two rounds of combat (the maximum) they are unable to force the Romans to retreat. Since the last round of battle has been fought, the Barbarian units that entered as Reserves are forced to retreat to an adjacent friendly area. Vercingetorix must retreat back into the fortress since he sortied out.}

\textit{With the field battle now complete, the Romans may proceed with their remaining two rounds of siege.}

\textbf{C. Crossing the Rhine Into Battle}
\blockquote[Book IV, Chapter 16]{With the German war concluded, I decided that I must cross the Rhine. Several reasons prompted me. The strongest was that I could see the Germans were all too ready to cross into Gaul, and I wanted them to have reasons of their own for anxiety when they realized that an army of the Roman people could and would cross the Rhine.}
\par
When the Romans or Germans cross the Rhine into Germania or a Gallic area the ensuing battle is restricted to a maximum of two rounds instead of the normal three.

This restriction only applies if the main group attacked across the Rhine. It does not apply if only reserves crossed the Rhine.

Every German unit killed in battle immediately scores 1 VP for the Roman player, except Ariovistus, which scores 2 VP for the Roman player.

\textbf{Germanic Strategic Withdrawal Option}
\par
The Barbarian player may strategically withdraw any or all German units in Germania, including Ariovistus, from the board before battle begins. This is not like a retreat, the units are not moved to a friendly area. Instead they are removed from the game.

Every German unit that strategically withdraws does not score any VP for the Roman player, but it is then permanently removed from the game.

Note that the strategic withdrawal option is only available for German units in Germania. It does not apply to German units fighting outside of Germania, and it can only be taken prior to battle. Once battle begins, the option to strategically withdraw is no longer available.

\textit{Designer's Note: This is designed to simulate the strategic value of Caesar's raids across the Rhine. Historically, these did not cause much in the way of actual damage or casualties because the Germanic tribes withdrew rather than risk a pitched battle. What it did accomplish, however, was to demonstrate to the Germans that the Romans could cross the Rhine at any time, and effectively convince them not to cross the Rhine into Gaul again.}
\par
\textit{Thus, the Barbarian player has to make a choice whether to risk giving the Roman player more VP's through a pitched battle, or to minimize the Roman VP potential at the cost of permanently losing control of German units.}

\textbf{D. The Alps}
\par
Units defending in the Alps (the Helvetii/Nantuates home area) receive double defense, i.e. it requires two hits in battle to score one point of damage. As with forts, if there is more than one unit defending, any half hits must be applied to a unit that has already suffered a half hit first.

\textbf{E. Amphibious Invasions}
\par
Barbarian units defending against Roman units that entered the area via naval movement are all treated as having one better initiative rating, i.e. any "C" units are considered to have a "B" initiative, and any "B" units are considered to have an "A" initiative. Any units already rated an "A" are unaffected.

In addition, there are only two rounds of battle instead of the normal three (and two rounds of siege instead of the normal three if a fort is present).

This only applies if the main force entered by naval movement. If a reserve force entered by sea, but the main force entered by land, then fight the battle normally.

\textbf{F. Retreat}
\par
Instead of firing, each unit may retreat to an adjacent friendly or empty area, or into a fortress if the area contains one and there is room for it. Units may retreat to the same or different areas. Retreating units are returned to their face up position before retreating, potentially concealing which units are retreating to any particular area. Border limits apply to retreating units on a per-battle-round basis, e.g. four units could retreat across a blue border, but no more than 2 per battle round

Units may not retreat into an area through which enemy units entered the area. However, if both players moved units across the same border in the same round, only the player who crossed the border last may retreat across that border.

Units may not retreat into a contested area, i.e. an area where there is an unfought battle. Only German units may retreat into Germania.

\textbf{Naval Retreat}
\par
Roman units conducting a naval move into an enemy port may retreat to any friendly port area in the same sea zone. They may not retreat if there is no friendly port area in the same sea zone.

Any units defending in a port area may retreat up to two units (maximum, not per battle round) to another friendly port area in the same sea zone.

\textbf{G. Regroup}
\par
The player left controlling the area at the end of battle is considered the victor and may regroup surviving units. Regroup allows all victorious units, including any in reserve, to move to any adjacent friendly or empty area.

A Unit cannot regroup into an unfought battle. Only German units may regroup into Germania. Border limits apply to regrouping units.

Roman units (only) in a port area may regroup into a friendly or empty port area in the same sea zone, up to a maximum of two units per area.