\clearpage
\section{Designer's Notes}
Differences between first and second editions of Caesar's Gallic War.

\subsection{Map}
\begin{itemize}
	\item There are two additional areas.
	\item There are now two sea zones instead of one.
\end{itemize}

\subsection{Units}
\begin{itemize}
	\item Greater variability between the units, both in terms of strength and initiative rating.
  	\item Areas that used to contain multiple blocks with the same tribe now have two distinct tribes, potentially with different values.
  	\item Some areas will start with different units at the start of each game.
 	\item The Helvetii are replaced by the Nantuates if eliminated. 
\end{itemize}
  
\subsection{Victory Points}
\begin{itemize}
	\item Now uses a cumulative VP system instead of a fixed VP level.
  	\item Yearly objectives added for flavor.
\end{itemize}
  
\subsection{Movement}
\begin{itemize}
  	\item The maximum number of units that can cross a black border has been reduced from six to four.
\end{itemize}
  
\subsection{Combat}
\begin{itemize}
  \item Now includes a fortress/siege system to replace the simple bonuses that existed before.
  \item Attacks across the Rhine, as well as amphibious invasions, are now restricted to two rounds of combat instead of the normal three.
  \item German units can now strategically withdraw (i.e. leave the game) instead of getting farmed for points.
\end{itemize}
  
\subsection{Supply}
\begin{itemize}
  \item Cards are now doubled for Romans (instead of value + 1).
  \item Romans use supply to recruit reinforcements instead of fixed arrivals.
  \item The cost to supply Roman legions in the field is now fixed at 1.
  \item The harvest roll now only affects garrison limits, it does not affect the cost to garrison.
\end{itemize}
  
\subsection{End of Turn}
\begin{itemize}
	\item German and Gallic replacements happen before eliminated units are returned to the map. This fixes a quirk of the old system where eliminated units would be back nearly to full strength right away.
\end{itemize}
  
\subsection{Solitaire}

A much more formal set of solitaire rules have been added.