\section{Victory Conditions}
\par
The Barbarian player wins instantly if Caesar is killed. Otherwise, victory is checked at the end of turn 8 (51 BC).

If the Roman player has scored at least 90 victory points, then the Roman player has scored a Minor Victory. Caesar returns to Rome, receives an Ovation and manages to avoid prosecution thanks to a modicum of popularity. However, his career wanes after that, and he fades into obscurity.

If the Roman player has scored at least 100 victory points, then the Roman player has scored a Major Victory. Caesar returns to Rome, receives a Triumph, 
and easily defeats prosecution attempts. Within a few years he leads an invasion against the Parthian Empire.

If the Roman player has scored at least 110 victory points, then the Roman player as scored a Massive Victory. History repeats itself. Caesar takes an army into Italy and the die is cast.

If the Roman player has scored less than 90 victory points then Barbarian player is considered victorious. Caesar returns to Rome without an army and is prosecuted for crimes during his consulship.

In addition to the cumulative end of turn VP scored, the following events also affect the Roman player's VP score:

\begin{itemize}
  \setlength\itemsep{0em}
  \item Each German non-leader unit eliminated: +1 VP
  \item Ariovistus eliminated: +2 VP
  \item Vercingetorix eliminated: +3 VP
  \item Amphibiously invade Britannia with at least 2 legions and fight at least one round of combat: +2 VP*
  \item Invade Germania with at least 2 legions and fight at least one round of combat and/or force at least one German unit to withdraw: +2 VP*
  \item Each Gallic home unit in Britannia eliminated: +1 VP (max 2 VP)*
  \item Each Roman legion unit eliminated: -5 VP
\end{itemize}

Note that the points for invading Britannia and Germania can only be scored once during the game. The number of VP for killing Gallic home units in Britannia is restricted to a maximum of 2 VP.

* Ignore this if playing with the optional Yearly Objectives.